### 2017-11-01 inst/documents/rotations1.Rmd
rescale migration 0-1, don't plot refuge if migration=0

---
title: "Rotations and Insecticide Resistance initial plots"
output:
  pdf_document:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, eval=TRUE, include=FALSE}

library(knitr)
opts_chunk$set(fig.height=7, fig.width=7)
# opts_chunk$set(dev="tiff",
#                dev.args=list(compression="lzw"),
#                dpi=300,
#                cache=FALSE,
#                fig.path='pap2figs/')
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
  library(rotations)
  library(ggplot2)
```

# Current state of play

1. Flexible number of insecticides between 1 & 5
1. Exposure can be set for each insecticide or same for all.
1. Starting frequency can be set for each insecticide or same for all.

Some example plots follow, these act as a visual test that the model is doing what we expect.

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="exposure set to 1 for r1 and 0.1 for r2"}

#expo_hi=c(1,0.1)
dfr <- run_rot(n=2, rotation_interval=100, expo_hi=c(1,0.1))

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotation interval 50, migration 0.01"}

# Interval 50, migration 0.01

# TODO remove hardcode fitnesses from here
dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.01)

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="decrease migration to 0.001"}

# Decrease migration to 0.001

dfr <- run_rot(rotation_interval=50, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, migration=0.001)

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="cost 0.1 effectiveness 1"}
dfr <- run_rot(n=3, rotation_interval=50, eff=1, cost=0.1)
```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Rotate when resistant, exposure 0.5 migration 0.01"}

### Rotate when resistant, migration 0.01

dfr <- run_rot(rotation_interval=0, max_generations = 300, hardcode_fitness = TRUE, same_insecticides =TRUE, expo_hi=0.5, migration=0.01)

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="Showing how with cost (0.1) an initial decline in the freq of the later insecticides when they are not being used can lead them to behave differently than earlier ones."}

run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.1 )

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.show='hold', fig.align='left', fig.cap="When cost is set to 0 all insecticides give the same curves."}

run_rot( n_insecticides = 3 , max_generations = 150 , start_freqs = 0.1 , rotation_interval = 50 , eff = 0.6 , dom = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0 )