### 2018-03-20 inst/documents/phaps-rotations-better.Rmd
thorough check

---
title: "perhaps rotations are better than sequences ?"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

  ```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(rotations)
library(ggplot2)

outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

resistance frequencies seem to take longer to rise for rotations than sequences once cost (or rufugia) are included. 
particularly when the rotation interval is short

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=9}

#for above
#\pagebreak

linputs <- list()

linputs$n_insecticides <- 2
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.1 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

linputs$rot_interval=0
dfseq <- do.call(run_rot, linputs)
linputs$rot_interval=6
dfrot <- do.call(run_rot, linputs)

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

```