### 2018-04-04 inst/documents/sensi_an_rotations1.Rmd
add extra outputs

---
title: "sensi_an_rotations1. Start of sensitivity analysis for rotations."
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
  library(rotations)
  library(ggplot2)
  library(gridExtra)

  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?

Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.

Test 2. Dominance of resistance and fitness costs both set to 0.5.
I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles. I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same. I’m probably wrong but it’s a nice test…..

Testing sensitvity analysis to answer question 1. Are sequences and rotations equivalent when coverage 100% and no fitness costs. On each page should see a run with 1 set of inputs first for sequence then for rotation.

\pagebreak

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}
#results='hide',
#comment=NA removes ## before text output

#start small & build it up
#maybe create an equiv. of resistance::sensiAnPaperPart.r later

# example for 1 run
# linputs <- list(max_gen = 50, n_insecticides = 2)
# do.call(run_rot, linputs)
# dfinputs <- as.data.frame(linputs)

# So one way of doing is to create an linputs list for each run & pass to run_rot()
# so run_rot() only ever runs for 1 scenario
# in contrast resistance::runModel2() accepted an input matrix & would run for multiple scenarios
# not sure which is better

n_scenarios <- 200

df_res_all <- data.frame()

for(i in 1:n_scenarios)
{
  # inputs (could also be saved to a dataframe first & extracted one row at a time)
  linputs <- list()

  # for Ians Q1 set coverage to 1 & cost to 0 
  # April 2018, having go at including variable cost, may want to do some cost=0 runs later

  ## constant inputs ##
  linputs$coverage <- 1
  linputs$migration <- 0
  #linputs$cost <- 0 #variable cost below
  linputs$max_gen <- 500

  #beware min_rwr_interval, it's to stop repeated small rotations
  linputs$min_rwr_interval = 5
  #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
  linputs$plot <-     FALSE

  ## variable inputs ##
  linputs$cost <-           runif(1, min=0,   max=0.1) 
  linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
  linputs$male_expo_prop <- runif(1, min=0,   max=1)
  linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
  linputs$rr  <-            runif(1, min=0.1, max=0.9)
  linputs$dom_sel <-        runif(1, min=0,   max=1)
  linputs$dom_cos <-        runif(1, min=0,   max=1) 
  linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
  linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1

  #random rot interval integer
  linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1

  # create null objects used plater for plotting
  ggrot <- ggseq <- NULL

  # to compare sequence to rotation run all the scenarios above for both seq & rot 
  # beware has to be 1:0 because interval is overwritten for seq
  for(rot_or_not in 1:0)
  {  
    #to get rot_int of 0 every other run
    if (rot_or_not == 0) linputs$rot_interval <- 0

    ##################
    # run one scenario
    dfres <- do.call(run_rot, linputs)

    # summarise results per scenario
    # NOTE this is also currently calc at end of run_rot() here or other post-process is probably better
    # and put in the output dfres, but the value has to be rpeated for every generation 
    res <- dfres %>%
      # only assess in control areas not in refugia  
      dplyr::filter(active_or_refuge=='active') %>%
      group_by(resist_gene) %>%
      # for all insecticides in all generations  
      # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
      # summarise(mean_gens_under50 = mean(gens_under50)) %>%
      # just for deployed insecticides 
      summarise(gens_dep_under50 = sum(resistance < 0.5 &
                                       #finds insecticide in use = this one
                                       resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
      summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
      unlist()

    # create dataframe from inputs and outputs
    # first add scenario column onto results
    res$scenario <- i
    df_in_out <- as.data.frame(c(linputs, res))

    # rbind this scenario inputs & results onto previous
    # not a good speed or memory way of doing ... but a start
    df_res_all <- rbind(df_res_all, df_in_out)

    ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)
    gens_res <- as.numeric(gens_under_thresh(dfres))

    # to plot results for checking
    if ( linputs$rot_interval == 0 )
    {
      ggseq <- ggres
      gens_seq <- gens_res
    } else
    {
      ggrot <- ggres
      gens_rot <- gens_res
    }

  } #end rot_or_not loop

  #printing input values
  #print(unlist(linputs))

  # try to plot input vals instead
  # extract out those between 0 & 1
  dfi <- data.frame(t(as.data.frame(linputs)))
  names(dfi) <- 'value'
  dfi$inputs <- rownames(dfi)

  #take out columns not between 0 & 1
  dfi <- filter(dfi, !inputs %in% c('max_gen','rot_interval','n_insecticides','plot','min_rwr_interval'))
  #also remove coverage & migration both set to 0 for now
  dfi <- filter(dfi, !inputs %in% c('coverage','migration'))

  ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
            geom_point() +
            theme_minimal() +
            ylim(0,1) +
            theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))

  grid.arrange(ggrot, ggseq, ggins, ncol=1)

  cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")

  #can I give the code to rerun this scenario
  cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
              "cost=",signif(linputs$cost,2),", ",
              "start_freqs=",signif(linputs$start_freqs,2),", ", 
              "rot_interval=",linputs$rot_interval,", ",
              "eff=",signif(linputs$eff,2),",\n",
              "dom_sel=",signif(linputs$dom_sel,2),", ",
              "dom_cos=",signif(linputs$dom_cos,2),", ",
              "rr=",signif(linputs$rr,2),", ",
              "expo_hi=",signif(linputs$expo_hi,2),", ",
              "coverage=",signif(linputs$coverage,2),",\n",
              "migration=",signif(linputs$migration,2),", ",
              "max_gen=",linputs$max_gen,", ", 
              "min_rwr_interval=",linputs$min_rwr_interval,            
              ")\n") )

} #end scenario loop

# save object containing inputs and results as rda for analysis
save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))

```

\pagebreak

# summary plot of all runs, would show any differences between sequence & rotations as non horizontal lines

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}

# want to plot line between sequence and rotation value
# any deviation from horizontal will be obvious
# could use rotation interval as the x axis, so that longer intervals will be clearer

# maybe not for publication
# but useful for quick view

    gg <- ggplot( df_res_all, aes_string(x='rot_interval',
                                         y='tot_gens_dep_under50',
                                         group='scenario',
                                         colour='scenario') ) +
          geom_line( alpha=0.5, lwd=1 )

    plot(gg)

# check proportion of scenarios not reaching thresholds by max_runs

    nruns_fail <- df_res_all %>% 
                  filter(tot_gens_dep_under50 == linputs$max_gen) %>%
                  summarise(n=n()) %>% 
                  unlist()

    cat('proportion runs not reaching thresholds : ', nruns_fail/(2*n_scenarios) ) #2* beacuse eeach scenario run for seq & rot

# plot gens againts exposure and effectiveness
# as a quick test to see if I can set the starting values of these higher

    gg <- ggplot( df_res_all, aes_string(x='eff',
                                         y='tot_gens_dep_under50') ) +
          geom_point( alpha=0.5 )
    plot(gg)

    gg <- ggplot( df_res_all, aes_string(x='expo_hi',
                                         y='tot_gens_dep_under50') ) +
          geom_point( alpha=0.5 )
    plot(gg)

    gg <- ggplot( df_res_all, aes_string(x='expo_hi*eff*male_expo_prop',
                                         y='tot_gens_dep_under50') ) +
          geom_point( alpha=0.5 )
    plot(gg)

```