### 2018-04-27 R/sensi_an_param_call_list.Rmd
starting paper3 and developing sensi analysis

---
  title: "sensi_an_param_call_list"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

To create a list of function calls one for each scenario in the sensitivity analysis.

To allow any run from the sensitivity analysis to be rerun and checked.

Relies on the seed being the same as in sensi_an_rotations.

A better option might be to read a line from df_res_all to extract the input params to rerun the model.

df_res_all[624,1:15]