### 2018-05-10 inst/documents/test-plot-resize.Rmd

---
title: "test plot resize"
# output:
#   pdf_document:
#   keep_tex: true
# fig_caption: true
# latex_engine: pdflatex

output: word_document
---

Testing setting width of plots for publication at 300 dpi into word doc.

Solution : just use fig.height & width (out.height etc not supported for word)
This makes the tiffs correct. Word foten resizes them. To get right in word, right click set size & position, just click reset to go back to 100%.

Was finding that plots ignore my size settings both in the tiff files and in the word doc.

```{r, eval=TRUE, include=FALSE}
#!!!! PAPER PROTOCOL
#!!!! submission MS, knit to word_document 
#!!!!  in WORD: delete figs, set double space, add line numbers 
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='')
```

```{r test-resize1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=2, fig.show='hold', fig.align='center', fig.cap="test"}

plot(1:10)
```

```{r test-resize2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=2, fig.show='hold', fig.align='center', fig.cap="test"}

plot(1:10)
```