### 2020-01-08 inst/documents/rotations_runs2020_sam.rmd
edit sams file for 160k runs to output tot_gens and mean mortality

---
title: "rotation_runs2020_sam"
author: "Andy South, Sam Jones and Ian Hastings"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

An rmarkdown document to run 160k runs, 10k from each of 16 experiments.
Sam Jones runs on fast machine overnight.
Output files are saved to the specified outFolder.

rotation_runs2019.Rmd is similar but allows experiments to be run individually and to generate a pdf of results for an experiment per run.

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(rotations)
library(ggplot2)
library(gridExtra)
library(readxl) # for reading the excel file that allows me to keep formatting to see inputs clearly

#Sam: Andy, it appears that TinyTex is also required 
library(tinytex)
outFolder <- "I:/Rotations/Results_Jan/"
```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5.2, comment=NA, results='asis'}

Stime = date ()  
cat("startTime", Stime, "\n")

plotyn <- 0 #allow turning plots on and off from here, may want to turn plots off if doing large number of runs

runList<-c('A1','A2','A3','A4','B1','B2','B3','B4','C1','C2','C3','C4','D1','D2','D3','D4')
#runList<-'A1'

nscenarios <- 10

for(r in runList){
  id_expt <- r

# # read an experiment file specifying ranges
# inex <- read_in_expt(infile, nscenarios=nscenarios) 
inex <- read_in_expt(nscenarios = nscenarios, multi=TRUE, id_expt=id_expt)

# create an object containing inputs for all scenarios
linmulti <- set_run_inputs(inex=inex)
# run all scenarios
for(i in 1:nscenarios)
{
  #message should output to the R markdown console
  if (i%%200 == 1) message("scenario ",i," of ",nscenarios," ",Sys.time())

  linputs <- get_one_in(linmulti, scen_num=i)

  if (linputs$rot_interval == 0) stop("in rotation_runs2019 rot_interval in input file needs to be > 0 to allow it to be compared to a sequence")

  linputs$plot <- plotyn #allows setting whether to plot from script
  linputs$add_gens_under50 <- TRUE #to add text to plots of gens under threshold

  #record rotation rot interval locally otherwise it gets overwritten by 0
  rot_rot_interval <- linputs$rot_interval

  # to compare sequence to rotation run all the scenarios above for both seq & rot 
  # need to do the rotation 2nd so that rot_interval from the input file can be saved with the outputs
  for(rot_or_not in 0:1)
  {  
    #to get rot_int of 0 in alternate runs - the other run uses the value from the input file
    if (rot_or_not == 0) linputs$rot_interval <- 0
    else linputs$rot_interval <- rot_rot_interval

    ##################
    # run one scenario
    dfres <- do.call(run_rot, linputs)

    # summarise results per scenario
    # this calcs for whole run, at end of run_rot() it calcs per insecticide
    if ( linputs$mort_or_freq == 'freq' ) threshold <- linputs$freq_thresh
    else if ( linputs$mort_or_freq == 'mort' ) threshold <- linputs$mort_thresh

    gens_u <- gens_under_thresh(dfres=dfres, threshold=threshold, mort_or_freq=linputs$mort_or_freq)

    # 10/12/19 change to recording end generations and mean mortality
    gens <- calc_max_gens(dfres=dfres)
    cum_mort <- cumulat_mortality(dfres=dfres)

    # to be able to save as for rot or seq
    if ( linputs$rot_interval == 0 ) {gens_seq <- as.numeric(gens)
                                      gens_u_seq <- as.numeric(gens_u)
                                      mort_seq <- cum_mort/gens_seq}
    else                             {gens_rot <- as.numeric(gens)
                                      gens_u_rot <- as.numeric(gens_u)
                                      mort_rot <- cum_mort/gens_rot}

    #code to rerun approx this scenario, can be copied from pdf and pasted into R
    #has to cope with converting vectors to strings
    #spaces between args stop the pdf from messing with words
    cat(paste0( "run_rot(rot_interval=",linputs$rot_interval,", ", 
                "n_insecticides=",linputs$n_insecticides,", ",
                #"mort_or_freq='",linputs$mort_or_freq,"', ", #quotes dont work, freq is now default
                #'mort_or_freq="',linputs$mort_or_freq,'", ',
                "start_freqs=",signif(linputs$start_freqs,2),", ", 
                "cost=c(",paste(signif(linputs$cost,2),collapse=','),"), ", 
                "eff=c(",paste(signif(linputs$eff,2),collapse=','),"), ",               
                "dom_sel=c(",paste(signif(linputs$dom_sel,2),collapse=','),"), ",
                "dom_cos=c(",paste(signif(linputs$dom_cos,2),collapse=','),"), ",
                "rr=c(",paste(signif(linputs$rr,2),collapse=','),"), ",
                "expo_hi=",signif(linputs$expo_hi,2),", ",
                "male_expo_prop=",signif(linputs$male_expo_prop,2),", ",
                "coverage=",signif(linputs$coverage,2),", ",
                "migration=",signif(linputs$migration,2),", ",
                "max_gen=",linputs$max_gen,", ", 
                "change_interval=",linputs$change_interval,
                #"min_rwr_interval=",linputs$min_rwr_interval,  
                ")\n\n") )

  } #end rot_or_not loop

  # add scenario id column and difference onto results
  # df_in_out is one row
  df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, 
                                        gens_u_rot=gens_u_rot, gens_u_seq=gens_u_seq,
                                        mort_rot=mort_rot, mort_seq=mort_seq, id=i))

  df_in_out$gens_rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq
  df_in_out$gens_u_rot_minus_seq <- df_in_out$gens_u_rot - df_in_out$gens_u_seq  
  df_in_out$mort_rot_minus_seq <- df_in_out$mort_rot - df_in_out$mort_seq

  # if the first scenario create df to store results of all scenarios
  if (i==1)
  {
    df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=nscenarios))
    names(df_res_all) <- names(df_in_out)
  }

  #putting results into one row of overall results dataframe
  df_res_all[i,] <- df_in_out

  #can save partial outputs in case doesn't make to end
  #if (i==nscenarios | i%%1000 == 0)
  #{

    # save object containing inputs and results as rda for analysis
    # by default saved to inst/documents folder

  #}

} #end scenario loop

save(df_res_all, file=paste0(outFolder,'df_res_all_rot',id_expt,nscenarios,'.rda')) #paste0(outFolder,'*.rda')) 
    # also save the experiment input file
    save(inex, file=paste0(outFolder,format(Sys.time(), "%Y-%m-%d-%H-%M-"),'inex-',id_expt,'.rda'))

Etime = date ()	
cat("endTime", Etime, "\n")

cat(Stime,Etime)
}
```