### 2020-12-14 inst/documents/combined_SI.tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Information: PRCC and Classification trees},
  pdfauthor={Andy South, Sam Jones and Ian Hastings},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Supplementary Information: PRCC and Classification trees}
\author{Andy South, Sam Jones and Ian Hastings}
\date{2020-05-19}

\begin{document}
\maketitle

The results shown on Figures 7 and 8 of the main text were subjected to
PRCC and decision tree analysis. To recap, the analysis was as follows
i.e.

Column A: each insecticide had identical properties; switching criterion
was resistance allele frequency exceeding 50\% Column B: each
insecticide had different properties; switching criterion was resistance
allele frequency exceeding 50\% Column C: each insecticide had identical
properties; switching criterion was mortality \textless{}90\% Column D:
each insecticide had different properties; switching criterion was
mortality \textless{}90\%

Row 1: no costs of resistance or dispersal into/out of an untreated
refugia\\
Row 2: costs are present; untreated areas absent\\
Row 3: costs absent; refugia present of an untreated refugia\\
Row 4: costs and refugia and both present\\

\pagebreak

PRCC

The results are given in the graphs at the end of this SI. The wealth of
data they contain meant the most convenient method of evaluating the
results is to print out each page, physically arrange them by column/row
and search for consistent patterns within rows or within columns. It is
also possible to compare consistency between evaluation criteria in each
row/column by comparing evaluation based on operational lifespan of the
insecticide armoury (red crosses) with evaluation based on mean
mortality during this lifespan (blue crosses); note that lifespan data
are absent in plots of A1 and C1 as each policy had identical results in
these circumstances.

There may be some slight patterns (e.g.~``coverage'' across row 3,
``effectiveness'' in column C) but these may well be coincidental and
there seem no clear, consistent patterns that would identify
circumstances under which one policy is consistently favoured over the
other.

The results given in D3 and D4 are the most important as they reflect
the most realistic deployment scenarios i.e.~insecticides differ in
their properties, switching is based on mortality), but PRCC values are
small and inconsistent both within the plots (i.e.~evaluation based on
operation lifespan or on mean mortality) or between analyses of D3 and
D4.

Finally, note the `p' value may be very low (the sample size is very
large) so the effects are `real' but small in magnitude and inconsistent
across scenarios

\pagebreak

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.279497e-05              4.132992e-09             2.839076e-271 
## Dominance of   resistance                  Exposure             Male exposure 
##             5.796504e-134              1.597436e-10              0.000000e+00 
##        Starting frequency 
##              1.962586e-28
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-2.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.431292e-10              6.866538e-02              7.066883e-02 
## Dominance of   resistance                  Exposure             Male exposure 
##             3.410289e-101              2.987057e-03              5.003460e-05 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              2.009100e-04              4.302566e-01             2.149941e-139 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##             6.256138e-103              6.819722e-58              8.500737e-01 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.204792e-72             1.070692e-190              1.688966e-70 
##        Starting frequency        Cost of resistance         Dominance of cost 
##             2.058805e-124             6.201080e-282              8.977118e-09
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-3.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.040190e-07              8.847019e-15              1.544334e-37 
## Dominance of   resistance                  Exposure             Male exposure 
##             1.590978e-248              3.102177e-38              5.230439e-10 
##        Starting frequency                  Coverage                 Dispersal 
##              2.436760e-98             3.112878e-287              2.048830e-04 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.736910e-28              2.438733e-12             4.530652e-206 
## Dominance of   resistance                  Exposure             Male exposure 
##              2.569953e-12              2.352183e-40              1.960913e-21 
##        Starting frequency                  Coverage                 Dispersal 
##              1.982413e-36              0.000000e+00              1.339645e-39
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-4.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              5.964994e-04              9.664349e-04              9.390170e-06 
## Dominance of   resistance                  Exposure             Male exposure 
##             1.232597e-103              1.023258e-13              8.119258e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.217907e-04              9.860850e-01              2.460829e-40 
##                  Coverage                 Dispersal 
##              5.473628e-25              2.552781e-01 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              3.420232e-47              1.749062e-30              1.830962e-38 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.030718e-35              2.399366e-60              5.004423e-23 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              7.936141e-59              4.081867e-44              9.055371e-11 
##                  Coverage                 Dispersal 
##              1.993505e-78              2.970499e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-5.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.527251e-51              8.903486e-01             2.042072e-319 
## Dominance of   resistance                  Exposure             Male exposure 
##             6.806321e-253              5.600780e-26              0.000000e+00 
##        Starting frequency 
##              4.181653e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-6.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##             3.445174e-107              2.530469e-04              5.221642e-06 
## Dominance of   resistance                  Exposure             Male exposure 
##              3.489091e-51              2.074492e-14              3.151796e-05 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.292085e-03              2.661099e-52              2.604509e-16 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.385269e-43              6.990974e-01              2.016125e-13 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.221363e-02              3.963614e-04              9.717417e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.914308e-04              3.310871e-13              3.074197e-06
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-7.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.621072e-44              1.190614e-03              3.178174e-11 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.321768e-59              4.243877e-35              5.891333e-06 
##        Starting frequency                  Coverage                 Dispersal 
##              1.081656e-78             7.213789e-200              1.513218e-64 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.922073e-25              2.194519e-01             4.420165e-108 
## Dominance of   resistance                  Exposure             Male exposure 
##              4.123880e-05              8.941555e-01              1.221618e-06 
##        Starting frequency                  Coverage                 Dispersal 
##              9.166678e-07             6.332364e-237              1.705509e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-8.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              3.091741e-52              9.523047e-03              5.946935e-11 
## Dominance of   resistance                  Exposure             Male exposure 
##              9.285600e-44              2.692993e-09              7.693886e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              3.210231e-04              1.814614e-25              4.636453e-06 
##                  Coverage                 Dispersal 
##              1.513121e-15              4.841748e-02 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.574257e-16              1.705323e-01              3.938262e-17 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.015848e-02              6.186243e-01              4.336015e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              3.418948e-04              1.734081e-05              1.140939e-07 
##                  Coverage                 Dispersal 
##              2.918941e-03              1.555878e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-9.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              5.536947e-04              0.000000e+00              0.000000e+00 
## Dominance of   resistance                  Exposure             Male exposure 
##              0.000000e+00             2.891056e-217             9.843181e-174 
##        Starting frequency 
##             1.976263e-323
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-10.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              6.613208e-02              9.126458e-01              2.107584e-03 
## Dominance of   resistance                  Exposure             Male exposure 
##              9.689983e-50              2.363406e-03              1.312496e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              2.947348e-03              7.551858e-01              1.364583e-63 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.901525e-24             3.162762e-177             1.467414e-309 
## Dominance of   resistance                  Exposure             Male exposure 
##              0.000000e+00             9.265694e-195              1.352644e-33 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              0.000000e+00              1.442660e-47              1.341460e-20
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-11.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.324957e-06              3.097366e-17              5.149826e-91 
## Dominance of   resistance                  Exposure             Male exposure 
##             6.974852e-181              2.268646e-56              1.139877e-05 
##        Starting frequency                  Coverage                 Dispersal 
##             9.857767e-129             1.725897e-225              2.580508e-05 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.962626e-02             2.666263e-161             2.505642e-144 
## Dominance of   resistance                  Exposure             Male exposure 
##             4.098326e-179              9.857408e-70              9.155748e-18 
##        Starting frequency                  Coverage                 Dispersal 
##             1.562930e-131              5.436548e-24              7.645491e-02
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-12.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.121770e-02              4.043657e-05              2.535163e-26 
## Dominance of   resistance                  Exposure             Male exposure 
##             7.777280e-145              5.690995e-33              3.447412e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              7.900528e-53              3.439374e-02              1.318700e-25 
##                  Coverage                 Dispersal 
##              2.240292e-66              1.772746e-02 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.831730e-05             2.018771e-175             1.446793e-136 
## Dominance of   resistance                  Exposure             Male exposure 
##             2.082715e-232             5.502352e-109              3.536926e-11 
##        Starting frequency        Cost of resistance         Dominance of cost 
##             4.439919e-225              3.034425e-04              1.179055e-07 
##                  Coverage                 Dispersal 
##              1.230631e-03              4.419448e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-13.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##                       NaN                       NaN                       NaN 
## Dominance of   resistance                  Exposure             Male exposure 
##                       NaN                       NaN                       NaN 
##        Starting frequency 
##                       NaN 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              6.242765e-52             1.633581e-122             1.296183e-124 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.561774e-79             7.049997e-117             9.553510e-153 
##        Starting frequency 
##             1.035562e-320
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-14.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              8.200503e-55              1.443596e-06              1.441594e-10 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.201341e-51              2.943605e-09              1.116616e-02 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              8.701257e-09              5.228213e-14              1.053013e-05 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              1.325812e-44              7.853500e-01              2.859244e-09 
## Dominance of   resistance                  Exposure             Male exposure 
##              1.031419e-05              2.370490e-08              8.829040e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.159202e-15              3.963494e-11              3.563846e-01
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-15.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.434045e-29              2.582068e-08              9.351446e-18 
## Dominance of   resistance                  Exposure             Male exposure 
##              6.528407e-38              6.660790e-23              1.426250e-05 
##        Starting frequency                  Coverage                 Dispersal 
##              9.760751e-81             1.531881e-125              9.664162e-29 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              4.154050e-15              2.993334e-32              5.412344e-13 
## Dominance of   resistance                  Exposure             Male exposure 
##              3.481640e-18              9.576041e-07              2.529119e-02 
##        Starting frequency                  Coverage                 Dispersal 
##              2.029387e-80              2.767899e-01              6.645464e-05
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/unnamed-chunk-3-16.pdf}

\begin{verbatim}
## [1] "P values for generations"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              7.191772e-48              2.713605e-14              8.983791e-21 
## Dominance of   resistance                  Exposure             Male exposure 
##              4.919087e-67              2.272904e-13              1.101868e-03 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              1.097601e-54              2.038884e-08              1.484036e-09 
##                  Coverage                 Dispersal 
##              2.970703e-27              6.020441e-01 
## [1] "P values for mortality"
##    Number of insecticides             Effectiveness    Resistance restoration 
##              2.143110e-29              3.130298e-02              1.966945e-03 
## Dominance of   resistance                  Exposure             Male exposure 
##              2.879619e-05              1.735534e-01              2.238551e-01 
##        Starting frequency        Cost of resistance         Dominance of cost 
##              4.868563e-15              8.983808e-07              6.133150e-01 
##                  Coverage                 Dispersal 
##              2.946694e-01              5.137393e-01
\end{verbatim}

\pagebreak

Classification trees

Classification trees were run on the same datasets, using the R package
Rpart. When evaluated on generations, those runs with a difference of 10
generations or less were regarded as ``equivalent''. A significant
subset of runs evaluated on mortality were also equivalent as they had
the same overall mortality under both policies (as might be expected if
both policies reached the same endpoint but by different routes). The
trees are interpreted as follows: • The numbers under each box give the
numbers of runs in the box which were equivalent, favoured rotations,
favours sequences, in that order. • The percentage given below each box
is the percentage of runs that fall into that box

As for PRCC, the only reasonable way to look for consistency seems to
require printing the graphs, spreading them on the floor in columns/rows
as in Figs 7\&8 in main ms and look for consistent patterns across row
and down columns. There was a tendency for low exposure to favour
rotations in Column A when evaluated on mean moratility but in general
there was no consistent patterns.

As noted above, scenario A3 and A4 reflect the most realistic deployment
scenarios but there was little consistence y between their trees.

Even when classification did occur they classes did not consistent
identify parameter spaces in which one policy was clearly favoured. For
example, Scenario D4 evaluated on mortality was relatively simple and
identified parameter spaces in which sequences, rotations or equivalence
was most likely. However, the numbers beneath each box show that
although one policy or equivalence tended to be favoured, there were a
large number of runs in the classification box which were contrary to
the favoured policy/equivalence. For example, rotations was the favoured
policy in the bottom right box of this figure with 134 run favouring
rotations but the classification box still contained 68 runs where the
policies were equivalent and 91 runs where sequences were favoured.

\pagebreak

\begin{verbatim}
[1] "Scenario A1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-1.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-2.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-3.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-4.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-5.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-6.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-7.pdf}

\begin{verbatim}
[1] "Scenario B1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-8.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-9.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-10.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-11.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-12.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-13.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-14.pdf}

\begin{verbatim}
[1] "Scenario C1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-15.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-16.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-17.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-18.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-19.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-20.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-21.pdf}

\begin{verbatim}
[1] "Scenario D1 Generations tree cannot run"
\end{verbatim}

\includegraphics{combined_SI_files/figure-latex/tree-22.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-23.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-24.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-25.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-26.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-27.pdf}
\includegraphics{combined_SI_files/figure-latex/tree-28.pdf}

\pagebreak

\end{document}