### 2020-12-14 inst/documents/paper_figs_rotations202010.Rmd

---
title: "paper_figs_rotations202010 : create paper figures from rotation runs"
#output: pdf_document
output: word_document
author: "Andy South, Sam Jones and Ian Hastings"
date: "`r Sys.Date()`"
---

Code to plot and analyse rotations runs for paper.

Runs themselves are done from rotation_runs2019.Rmd.

16 expts of 10,000 runs each comparing rotations and sequences.

```{r setup, eval=TRUE, include=FALSE}
#set to eval=TRUE to make figs for submission

library(knitr)

# options to create final publication quality figures
# word doc summarises & figs get stored
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='rotfigs/',
               fig.width=8.5,
               fig.height=8.5)

  library(rotations)
  library(ggplot2)
  library(gridExtra)
  library(tidyverse)
  library(patchwork) #for arranging multiple plots on page

  # where the rotation runs saved
  #infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
  #infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\SJ_Runs_Nov\\"
  infolder <- "C:\\Dropbox\\2017_Rotations for IRM\\2020-02-results-sam\\"

```

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}

# sam is running all 16 scenarios with a single outfile
# but in the meantime can I read the rda's in individually ?
# e.g. need to be able to find files named like this
# 2019-11-05-10-42-df_res_all_rotA210000.rda
# how to get all combinations of A-D and 1-4
id_columns <- expand.grid(c("A","B","C","D"),c(1:4))
expt_ids <- paste0(id_columns[,1], id_columns[,2])
#i think that if I add the 10000 in I can find the unique files
expt_ids <- paste0(expt_ids,10000)

#list all files
#list.files(infolder)

#list files matching a pattern, now get just on matching each
#list.files(infolder,"A110000")

df_res_16 <- data.frame(stringsAsFactors=FALSE)

# read in all 16 files, rbind and add ID columns
for(i in 1:length(expt_ids))
{

  # gets loaded as df_res_all because that is what it was saved as
  load(paste0(infolder, list.files(infolder, expt_ids[i])))
  # add experiment ids, first as separate columns to make facetting easier maybe
  df_res_all$idAD <- substr(expt_ids[i],1,1)
  df_res_all$id14 <- substr(expt_ids[i],2,2)

  df_res_16 <- rbind(df_res_16, df_res_all)

}

save(df_res_16, file=paste0(infolder,'df_res_16.rda'))

# filter just the 4 A expts
df_res_A <- dplyr::filter(df_res_16, idAD=="A")

```

#fig1 before this is model description

# example scenarios
# A1 frequency switch insecticides same, no costs no disp

```{r fig4-examp-run-A1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#2020-10 was fig2

dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )
ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )

ggseq <- ggseq + theme(axis.title.x = element_blank()) +
                 ggtitle("A) sequence")
ggrot <- ggrot + theme(legend.position="none")+
                 ggtitle("B) rotation")

#using patchwork
#plot seq and rot side-by-side
#guides=collect to put legend between plots, doesn't seem to work
(ggseq / ggrot) + plot_layout(guides = 'collect')

```

# A2 frequency switch insecticides same, costs no disp

```{r fig5-examp-run-A2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#2020-10 was fig3

dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )
ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE  )

ggseq <- ggseq + theme(axis.title.x = element_blank()) +
                 ggtitle("A) sequence")
ggrot <- ggrot + theme(legend.position="none")+
                 ggtitle("B) rotation")

#using patchwork
#plot seq and rot side-by-side
#guides=collect to put legend between plots, doesn't seem to work
(ggseq / ggrot) + plot_layout(guides = 'collect')

```

# C2 mortality switch insecticides same, no costs, disp

```{r fig6-examp-run-C2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

# mortality switch

#2020-10 was fig4

dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'mort')

dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0.02), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=1, migration=0,
min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'mort')

ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE, legend_title = '' )
ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=TRUE, plot_refuge = FALSE, logy =FALSE  )

ggseq <- ggseq + theme(axis.title.x = element_blank()) +
                 ggtitle("A) sequence")
ggrot <- ggrot + theme(legend.position="none")+
                 ggtitle("B) rotation")

#using patchwork
#plot seq and rot side-by-side
#guides=collect to put legend between plots, doesn't seem to work
(ggseq / ggrot) + plot_layout(guides = 'collect')

```

# A3 frequency switch insecticides same, no costs, disp

```{r fig7-examp-run-A3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#2020-10 was fig5

dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.001, cost=c(0), eff=c(0.98), dom_sel=c(0.96), dom_cos=c(0.32), rr=c(0.12), expo_hi=0.75, male_expo_prop=0.34, coverage=0.3, migration=0.1, min_rwr_interval=10, plot=FALSE, max_gen=500, mort_or_freq = 'freq')

ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, logy =FALSE )
ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, logy =FALSE )

ggseq <- ggseq + theme(axis.title.x = element_blank()) +
                 ggtitle("A) sequence")
ggrot <- ggrot + theme(legend.position="none")+
                 ggtitle("B) rotation")

#using patchwork
#plot seq and rot side-by-side
#guides=collect to put legend between plots, doesn't seem to work
(ggseq / ggrot) + plot_layout(guides = 'collect')

```

# B2 frequency switch insecticides different, costs

```{r fig8-examp-run-B2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#2020-10 was fig6

#in sequence first insecticide keeps being used for 500 gen
#in rotation switch to other insecticides that reach threshold means that generations are lost over threshold

dfseq <- run_rot(rot_interval=0, n_insecticides=3, start_freqs=0.0059, cost=c(0.099,0.088,0.062), eff=c(1,0.96,0.78), dom_sel=c(0.63,0.95,0.42), dom_cos=c(0.6,0.49,0.14), rr=c(0.36,0.11,0.86), expo_hi=0.65, male_expo_prop=0.087, coverage=1, migration=0, plot=FALSE, max_gen=500, change_interval=10, mort_or_freq = 'freq')

dfrot <- run_rot(rot_interval=10, n_insecticides=3, start_freqs=0.0059, cost=c(0.099,0.088,0.062), eff=c(1,0.96,0.78), dom_sel=c(0.63,0.95,0.42), dom_cos=c(0.6,0.49,0.14), rr=c(0.36,0.11,0.86), expo_hi=0.65, male_expo_prop=0.087, coverage=1, migration=0, plot=FALSE, max_gen=500, change_interval=10, mort_or_freq = 'freq')

ggseq <- rot_plot_resistance(dfseq, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )
ggrot <- rot_plot_resistance(dfrot, plot=FALSE, plot_mort=FALSE, plot_refuge = FALSE, logy =FALSE )

ggseq <- ggseq + theme(axis.title.x = element_blank()) +
                 ggtitle("A) sequence")
ggrot <- ggrot + theme(legend.position="none")+
                 ggtitle("B) rotation")

#using patchwork
#plot seq and rot side-by-side
#guides=collect to put legend between plots, doesn't seem to work
(ggseq / ggrot) + plot_layout(guides = 'collect')

```

# difference between rotations and sequences, all runs

```{r fig-si-1-abs-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#GOOD setting colours so can see easier when rot or seq better

widthbin <- 10 #either 10 or 20 looks fairly similar, 5 gives gaps because of course runs are constrained to 1 year

# df_res_all$best_strategy <- ifelse(df_res_all$gens_rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_all$gens_rot_minus_seq > (widthbin/2), "rotation", "no_difference"))
# 
# ggplot(df_res_all, aes(x=gens_rot_minus_seq, fill=best_strategy)) + 
#      geom_histogram(binwidth=widthbin, colour='white') + 
#      theme_minimal() +
#      theme(panel.grid.minor.y=element_blank())

# calc percentage difference between rotation and sequence for resistance & mortality
df_res_16$rms_percent <- 100 * df_res_16$gens_rot_minus_seq / df_res_16$gens_seq
df_res_16$rms_mort_percent <- 100 * df_res_16$mort_rot_minus_seq / df_res_16$mort_seq
#seems problem with above try Ians way
df_res_16$rms_mort_percent <- 100 * (df_res_16$mort_rot / df_res_16$mort_seq -1)
# Ian said this is equivalent and it does give nearly same result 10pow-14 difference
#df_res_16$ian_rms_percent <- 100 * (df_res_16$gens_rot / df_res_16$gens_seq -1)
#df_res_16$tst <- df_res_16$ian_rms_percent-df_res_16$rms_percent

# 2020-02-25 checking B1 scenarios that are giving weird results
df_res_B1 <- dplyr::filter(df_res_16, idAD=='B' & id14 == 1)

df_res_B1 %>% group_by(gens_rot_minus_seq) %>% tally()
# A tibble: 7 x 2
#   gens_rot_minus_seq     n
#                <dbl> <int>
# 1                -10     1
# 2                  0  6454
# 3                 10  1653
# 4                 20  1207
# 5                 30   514
# 6                 40   150
# 7                 50    21
# so this seems to suggest that rotations are indeed better for B1

#possibilities :
#A something different in the runs that Sam did and the 200 example ones done for the document
#B a problem with the document

#but the B1 document looks as expected ???

# TODO 2020-02-25 check whats going on here

# best strategy calculation
# no_difference set to +/- half of widthbin set above
df_res_16$best_strategy <- ifelse(df_res_16$gens_rot_minus_seq <= -(widthbin/2),"sequence", ifelse(df_res_16$gens_rot_minus_seq > (widthbin/2), "rotation", "no_difference"))

# OR use a % threshold for strategies to be the same
# still uses widthbin which now represents a pcent
# no_difference set to +/- half of widthbin% set above
df_res_16$best_strategy <- ifelse(df_res_16$rms_percent <= -(widthbin/2),"sequence", ifelse(df_res_16$rms_percent > (widthbin/2), "rotation", "no_difference"))

# lengthening facet labels
# idAD_lookup <- c(
#   A = "A freq i-same",
#   B = "B freq i-diff",
#   C = "C mort i-same",
#   D = "B mort i-diff"
# )
# id14_lookup <- c(
#   "1" = "1 nocost nodisp",
#   "2" = "2 costs nodisp",
#   "3" = "3 nocost disp",
#   "4" = "4 costs disp"
# )

idAD_lookup <- c(
  A = "A frequency switch\n insecticides same",
  B = "B frequency switch\n insecticides different",
  C = "C mortality switch\n insecticides same",
  D = "D mortality switch\n insecticides different"
)

id14_lookup <- c(
  "1" = "1 no fit. costs\n no dispersal",
  "2" = "2 fit. costs\n no dispersal",
  "3" = "3 no fit. costs\n dispersal",
  "4" = "4 fit. costs and\n dispersal"
)

#colrs <- c("no_difference"=rgb(210,210,210,max=255),"rotation"=rgb(113,166,89,max=255),"sequence"=rgb(137,117,202,max=255))
colrs <- c("no_difference"=rgb(255,255,255,max=255),"rotation"=rgb(2,98,215,max=255),"sequence"=rgb(227,0,169,max=255))

#gens_rot_minus_seq, facet by A-D and 1-4, better facet labels
ggplot(df_res_16, aes(x=gens_rot_minus_seq, fill=best_strategy)) + 
     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
     scale_fill_manual(values=colrs) +
     theme_minimal() +
     xlim(-100,100) +
     xlab("advantage of rotations over sequence in generations") +
     ylab("count of randomly sampled runs") +  
     theme(panel.grid.minor=element_blank(),
           legend.position="bottom") +
     facet_grid(rows=vars(id14), cols=vars(idAD),
                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

#tally
#df_res_16 %>% group_by(id12) %>% tally()

```

# pcent difference between rotations and sequences, all runs

```{r fig2-pcent-diff-rot-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

# 2020-10 previously fig7

#rms (gens_rot_minus_seq)_pcent, facet by A-D and 1-4, better facet labels
ggplot(df_res_16, aes(x=rms_percent, fill=best_strategy)) + 
     geom_histogram(binwidth=widthbin, colour='darkgrey') + 
     scale_fill_manual(values=colrs) +
     theme_minimal() +
     #xlim(-100,100) +
     scale_x_continuous(limits=c(-80,80), breaks = c(-60,-40,-20,0,20,40,60)) +
     xlab("percentage advantage of rotations over sequence in generations") +
     ylab("count of randomly sampled runs") +    
     theme(panel.grid.minor=element_blank(),
           legend.position="bottom") +
     facet_grid(rows=vars(id14), cols=vars(idAD),
                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

#just as a test
#can I plot gens_seq and gens_rot in same facet
#but I need to make data longer to do that, currently in same columns

```

```{r fig3-mort-which-best-meanmort, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

# 2020-10 previously fig8

# ** using best strategy from generations calc above
# ** so trying to see if being better in generations terms
# ** has consequences for mean mortality

#2020-10-08 Ian asked to change this to percent difference

ggplot(df_res_16, aes(x=rms_mort_percent, fill=best_strategy)) + 
#ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
#ggplot(df_res_16, aes(x=mort_rot_minus_seq)) +   
     #geom_histogram(binwidth=widthbin, colour='darkgrey') + 
     geom_histogram(binwidth=5, colour='darkgrey') + 
     scale_fill_manual(values=colrs) +
     theme_minimal() +
     xlim(-40,40) +
     xlab("percentage advantage of rotations over sequence in mean mortality") +  
     ylab("count of randomly sampled runs") +  
     theme(panel.grid.minor=element_blank(),
           legend.position="bottom") +
     facet_grid(rows=vars(id14), cols=vars(idAD),
                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

#old version just mortality rather than %
# ggplot(df_res_16, aes(x=mort_rot_minus_seq, fill=best_strategy)) + 
#      geom_histogram(binwidth=0.02, colour='darkgrey') + 
#      scale_fill_manual(values=colrs) +
#      theme_minimal() +
#      xlim(-0.2,0.2) +
#      xlab("advantage of rotations over sequence in mean mortality") +
#      ylab("count of randomly sampled runs") +  
#      theme(panel.grid.minor=element_blank(),
#            legend.position="bottom") +
#      facet_grid(rows=vars(id14), cols=vars(idAD),
#                 labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

```

# at each generations below threshold for rotations, which strategy was best

```{r fig9-which-best-each-gens-rot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

# 2020-10 fig9

#oooo this is interesting
#it shows what gens_rot the best strategies occur at

# but difficult to see the ones where there is an advantage to either strategy because so few of them 
# increasing binwidth to 50 made advantage run more visible

#also good that it shows runs lasting longer when fitness costs and dispersal are added

ggplot(df_res_16, aes(x=gens_rot, fill=best_strategy)) + 
     geom_histogram(binwidth=50, colour='darkgrey') + 
     scale_fill_manual(values=colrs) +
     theme_minimal() +
     #scale_y_log10() + #log scale did make it easier to see lower values but misleading
     #xlim(-100,100) +
     xlab("generations to reach thresholds for ROTATION") +
     ylab("count of randomly sampled runs") +    
     theme(panel.grid.minor=element_blank(),
           legend.position="bottom") +
     facet_grid(rows=vars(id14), cols=vars(idAD),
                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

#Ian asked if could make the bars mean of rot & seq
#but I think this is confusing & potentially wrong
# ggplot(df_res_16, aes(x=0.5*(gens_rot+gens_seq), fill=best_strategy)) + 
#      geom_histogram(binwidth=50, colour='darkgrey') + 
#      scale_fill_manual(values=colrs) +
#      theme_minimal() +
#      #scale_y_log10() + #log scale did make it easier to see lower values but misleading
#      #xlim(-100,100) +
#      xlab("generations to reach thresholds for ROTATION") +
#      ylab("count of randomly sampled runs") +    
#      theme(panel.grid.minor=element_blank(),
#            legend.position="bottom") +
#      facet_grid(rows=vars(id14), cols=vars(idAD),
#                 labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

```

# at each generations below threshold for rotations, which strategy was best

```{r fig9seq-which-best-each-gens-seq, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#same as previous but for seq rather than rot
#maybe to go in SI

#gens_seq the best strategies occur at
ggplot(df_res_16, aes(x=gens_seq, fill=best_strategy)) + 
     geom_histogram(binwidth=50, colour='darkgrey') + 
     scale_fill_manual(values=colrs) +
     theme_minimal() +
     #xlim(-100,100) +
     xlab("generations to reach thresholds for SEQUENCE") +
     ylab("count of randomly sampled runs") +    
     theme(panel.grid.minor=element_blank(),
           legend.position="bottom") +
     facet_grid(rows=vars(id14), cols=vars(idAD),
                labeller=labeller(idAD=idAD_lookup, id14=id14_lookup))

```

# count proportion of scenarios not reaching resistance thresholds

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# check proportion of scenarios not reaching thresholds by max_runs for both seq & rot

    nruns_fail <- df_res_16 %>% 
                  filter(gens_seq == max_gen & gens_rot == max_gen) %>%
                  summarise(n=n()) %>% 
                  unlist()

    cat('proportion runs not reaching thresholds : ', nruns_fail/nrow(df_res_16) )

    # remove runs where thresholds not reached for seq & rot 
    df_res_16_threshreached <- df_res_16 %>% 
                  filter(gens_seq != max_gen | gens_rot != max_gen)

    nruns_same <- df_res_16_threshreached %>% 
                  filter(gens_seq == gens_rot) %>%
                  summarise(n=n()) %>% 
                  unlist()

    cat('proportion runs (that did reach threshold) where rotations and sequence the same  : ', nruns_same/nrow(df_res_16) )

    nruns_rot_20pc_better <- df_res_16_threshreached %>% 
                  filter(gens_rot_minus_seq/gens_seq >= 0.2 ) %>%
                  summarise(n=n()) %>% 
                  unlist()

    cat('proportion runs (that did reach threshold) where rotations >20 % better than sequence  : ', nruns_rot_20pc_better/nrow(df_res_16) )

```